On the formation of H$\alpha$ around classical T Tauri stars

### 2006-01-11 TTau_Halpha/tmp.tex
Corrections made after Tim's comments. The subsection on the fits of T Tau is eliminated from kurosawa.tex.  The subsection was saved in a separate tex file: ttau_model_subsec.tex.  (RK)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@616 25464f21-6099-402c-b30a-640f1d40d38b

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amssymb}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{lscape}
\usepackage{times}

@@

We investigate the formation of H$\alpha$ from classical T~Tauri
stars with a complex circumstellar geometry -- the combination of
mangnetisherical accretion, collimated steller wind, and accretion
disc. AND MORE

@@

\begin{keywords}
stars:formation -- stars: individual: SU Aur -- circumstellar matter -- infrared: stars -- stars: pre-main-sequence
\end{keywords}

@@

T~Tauri stars (TTS) are young ($<\sim3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass stars, and known as progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some
of the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit an excess
amount of continuum flux in the ultraviolet (UV) and infrared (IR).
Their spectral energy distribution and polarization data suggest the
presence of a circumstellar disc, and it plays an important role in
regulating dynamics of gas flows around CTTS.

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profles show that the evidence for both outward wind
flows and inward accretion flows, as seen in the blueshifted absorption
features in H$\alpha$ profiles and the redshifted inverse P~Cygni
(IPC) profiles. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).

Recent H$\alpha$ spectro-astrometric observations by \citet{takami:2003}
show the direct evidence for the presence of bipolar and monopolar
outflows down to $\sim1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observation of high-resolution ($\mathrm{R=50\,000}$)
two-dimensional spectral of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended (SCALE?)
H$\alpha$ emission in the direction perpendicular to the obscuring
circumstellar disc, and in both above and below the disc --- suggesting
the presence of the bipolar outflows. In silightly larger scale, \emph{HST}
observation of HH30 (Burrorws et al 1996) shows the jet traced to
within $\lesssim30$~AU from the star. The jet has a cone shape with
an opening angle of $3^{\circ}$ between 70 and 700 AU. (Originally
from Konig and Pudritz 2000 review in PP IV). \citealt{alencar:2000}
found about 80 per cent of their samples (30 CTTS) show blueshifted
absorption componets in at least one of Balmer lines and Ca~K (most
commonly in H$\alpha$).

In a currently favoured model of accretion flows around CTTS, the
accretion discs are disrupted by the magnetosphere of stars which
channels the gas from the disc onto the surface of the stars (e.g.,
\citealt{uchida:1985}; \citealt{koenigl:1991}; \citealt{cameron:1993}\citealt{shu:1994}).
This picture of the accretion flows is supported by the evidences
that CTTS have relatively strong ($\sim10^{3}\mathrm{G}$) magnetic
field (e.g., \citealt*{jonhs-krull:1999}; \citealt{guenther:1996})
and by the radiative transfer models which reproduce the observed
profiles for some TTS (\citealt{muzerolle:2001}). The magnetospherical
accretion model naturally explains the blueward asymmetic emission
lines (seen in some of CTTS) caused by the partial occulation of the
flow by the disc, and the redshifted absorption componet at the typical
free-fall velocities (a few hundred $\mathrm{km\, s^{-1}}$) seen
in some of CTTS.

Dispite the success of the magnetosherical accretion model in explaining
the line profiles in some CTTS, the overwhelming observational evidences
for the outflow (mentoned above) in the CTTS profiles suggests that
this model is only a part of a whole picture. Clearly, the modefification
to include the outflowing wind/jet flow is neccessary if one requires
to predict the mass-accretion rate and the mass-loss rate of CTTS
by modelling their emission profiles (e.g. H$\alpha$).

Priory to the magnetospherical models, many alternative models had
been considered to explain the observed spectroscopic features mentioned
earlier. For example, 1.~the Alfv\'{e}n wave-driven wind model (e.g.~\citealt{decampli:1981};
\citealt{hartmann:1982}), 2.~turbulant boundary layer (between the
accretion disc and stellar surface) model (e.g.~\citealt{bertout:1988};
\citealt{basri:1989}), 3.~chromospheric model (e.g.~\citealt{calvet:1984}),
and 4.~disc wind model (e.g. \citealt{calvet:1992b}; \citealt{kwan:1995}).
Considering the sucess of the magnetospherical accretion flow model
in some cases and the observational eividences for the outflows, it
is likely that an improved spectroscopic model requires both inflow
and outflow components. The combination of the magnetospheric accretion
flow model with model 1 or 4 would be a reasonable starting point
for an improved spectroscopic model. Although time consuming to explore
a larger parameter space, more realistic density, velocity and temperature
structure from magnetohydrodynamic wind+accretion models should be
considered as inputs for an radiative transfer model in the future.

The magnetocentrifugal wind model, first proposed by \citet{blandford:1982},
has been often used to reproduce the large-scale wind structure of
T Tauri stars, or to model observed optical jets (e.g.~HH~30 jet
by \citealt{burrows:1996}; \citealt{ray:1996}). The lanching of
the wind from a Keplerian disc is typically done by treating the equaotorial
plane of the disc as a mass-injecting boundary condition (e.g., \citealt{ustyugova:1995};
\citet{shu:1995}; \citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).
Depending on the location of the open magnetic fields anchored to
the disc, two different types of winds are preoduced. If the field
is constrained to be near the co-rotation radius of stellar magnetosphere,
an {}``X-wind'' \citep{shu:1994} is produces. If the open field
lines are located in a wider area of the disc, a {}``disc-wind''
similar to \citet{koenigl:2000} is produced \citet{krasnopolsky:2003}.
Recent reviews on the jet/wind-disc connection can be found in \citet{koenigl:2000}
and \citet{pudritz:2005}.

In this paper, we present a semi-emperical wind -- accretion disc
-- magnetospherical accretion hybrid model for CTTS, and perform parametric
studies of the H$\alpha$ formation. This study should provide preliminary
physical conditions which lead to the wide varaiety of emission line
profiles seen in the observation, and would help to construct more
comprehensive circumstellar models (e.g. via MHD simulations) of T~Tauri
stars. the Using the model results, we examine the H$\alpha$ spectroscopic
classification used by \citet{reipurth:1996}, and compare models
with observations. We will also discuss whether our model is consistent
with some predictions made by the recentt MHD studies (e.g.~ $\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$).

In section~\ref{sec:model-configuration}, the model assumptions,
and the basic model configurations are presented. We discuss the radiaive
transfer model used to compute the profiles in section~\ref{sec:radiative-transfer-model},
and the results of model calculations are given in section~\ref{sec:Results}.
The closer examination and discussion of the results are presented
in section~\ref{sec:Discussion}. Finally, the summary of this work
and the conclusion are in section~\ref{sec:Conclusions}.

@@

To understand how the different part of the CTTS circumsterllar enviroment
contributes to the fomation of H$\alpha$, the model spaced is devided
into four different regions: 1.~Central contiuum source, 2.~magnetosphere
accretion flow, 3.~wind outflow, and 4.~accretion disc. Fig.~\ref{fig:overview}
depicts the relative location in the model space. In all regions,
the density is symmetric around the z-axis. The innermost raius of
the magnetsphere at the equotrial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas freely fall, moving along the magneic filed onto the
surface of the star. For the purpose of computational simplicity,
the colliminated wind are defined only in the outside of the largest
radius of the magnesophere. In the following subsections, the details
of model components will be described.

@@

\includegraphics[%
  scale=0.6]{figures/overview2.eps}

@@

\caption{Basic Model Configuration. The system consist of four components: 1.~the continuum source located at the origin $\left( o \right)$ of the cartisian coordinates $\left(x,y,z\right)$ -- the y-axis are into the paper, 2.~magnetospherical accretion flow, 3.~colliminated wind/jet outflow, and 4.~accretion disc. The density distribution is symmetric around the z-axis. The innermost raius of the magnetsphere (at the equotrial plane) coincides with the the inner radius of the accretion disc. From the inner most part of the accretion disc, the gas freely fall, moving along the magneic filed onto the surface of the star. For simplicit, the colliminated wind/jet exists outside of the largest radius of the magnesophere (dotted line). }

\label{fig:overview}

@@

\subsection{Contiuum Source}

\label{sub:Contiuum-Source}

Unless specified otherwise, we adopt stellar parameters of a typical
classical T~Tauri star (REF) for the central conitnuum source, i.e.
radius ($R_{*}$), mass ($M_{*}$), and effective temperature the
photosphere ($T_{\mathrm{ph}}$) are $2\, R_{\sun}$, $0.5\, M_{\sun}$,
and $4000\,\mathrm{K}$ respectively. The model atomsphere of \citet{kurucz:1979}
equivallently \citet{kurucz:1993}  with $T_{\mathrm{ph}}=4000\,\mathrm{K}$
and $\log g_{*}=3.5$ (cgs) is used to continuum flux at H$\alpha$wavelength.
The parameters are summarised in Table~\ref{tab:std_parameters}.

The limb-darkening law used for the continuum source is $I\left(\mu\right)=I_{0}\left(1+b\mu\right)$
where $\mu=\cos\theta$ and $\theta$ is the polar angle. $b$ is
a parameter. For no limb-darkening case, $b=0$, and for the Eddington
approximation case $b=3/2$. The former is used in all the calulcations
presented here. (Is this an ok assumption for T Tauri's? You need
to check this.)

@@

\begin{tabular}{lcccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$\beta$&
$v_{\infty}$&
$m$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[-]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{\sun}\right]$&
$\left[\mathrm{AU}\right]$\tabularnewline
\hline
Standard&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$0.2$&
$200$&
$2.0$&
$2.0$&
$100$\tabularnewline
\hline
\end{tabular}

@@

We adopt the magnetospherical accretion flow model of \citet{hartmann:1994},
as done so by \citet{muzerolle:2001} and by \citet{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:overview}, the innermost radius ($R_{\mathrm{mi}}$)
of the magnetophere at the equatorial plane ($z=0$) is assigned to
be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equaotrial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The former value corresponds to the corotation radius
of the accretion disc, and the geometry of the magnetic field/stream
lines is kept constant throghout this paper. The geometry of the magnetosphere
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form: $r=R_{\mathrm{m}}\sin^{2}\theta$ \citep[see][]{ghosh:1977}
where $r$, and $\theta$ are coordinates of the field point ($p$)
in Fig.~\ref{fig:overview} in spheical coodinates, and $R_{\mathrm{m}}$is
the radial distance to the field line at the equotorial plane ($\theta=\pi/2$).
The range of $R_{\mathrm{m}}$is restricted between $R_{\mathrm{mi}}$
and $R_{\mathrm{mo}}$ as mentioned above. Using the field geometry
above, the conservation of energy, and igoring the rotation of magnetosphere,
the velocity and the density of the accreting gas along the steam
line can be found. See \citet{hartmann:1994} for details.

In all of our models, the temperature structure of the magnetospheric
used by \citet{hartmann:1994} is also adoped here. The temperature
structure was computed assuming a volumetric heating rate which is
proportional to $r^{-3}$, and using the energy balance of the radiative
cooling rate of \citet{hartmann:1982} and the heating rate \citep{hartmann:1994}.
Although the temperature structure of the accretion steam could significantly
affect the line source function, for the purpose of the expolaring
the general characteristics of the H$\alpha$ formation, this simple
form is a reasonable assumption.\citet{martin:1996} presented the
self-consistent determination of the thermal structure of the inflowing
gas along the dipole magnetic field, same as the one used here, by
solving the heat equation couple to trate equations for hydrogen.
He found that main heat source is adiabatic compression due to the
convering nature of the flow, and the major contirbutors of the cooling
process are bremsstrahling radiation and line emission from Ca~II
and Mg~II ions. The temerature sturcture of \citet{martin:1996}
qualitaively agrees with that of \citep{hartmann:1994}. The sensitivty
tests of H$\alpha$ on the temperature structure will be presented
later (SPECIFY THE SECTION LATER).

As the infalling gas becomes closer to the stellar surface, it is
decelerates in a strong shock, and heated to $\sim10^{6}\,\mathrm{K}$
(with typical parameters). The X-ray radiation produced in the shock
will be absorbed by the gas located near, and reemitted in optical
and UV radiation (\citealt{koenigl:1991}; \citealt{hartmann:1994}).
This will create hot rings near the stellar surface where the magnetic
field intersects with the surface. For simplicity, the free-falling
kinetic energy is assumed to be thermalized in the radiating layer,
and reemits as blackbody radiation (with a single temerature). With
the parameters of the magnetosphere and the star given above (Table~\ref{tab:std_parameters}),
about $8$~per~cent of the surface is covered by the hot rings,
and its temperature is about $6400\,\mathrm{K}$. If the mass-accretion
rate is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, the ratio of this
accretion luminosity to the photospheric luminosity is about $0.5$.
The contiuum emission from the hot rings is taken into account when
computing the line profiles.

\subsection{Wind}

@@

\subsubsection{Colliminated wind model}

JUSTTIFICATION OF USING THIS FORMULATION. BREIFLY MENTIONED THE WORK
DONE BY OTHERS.

Based on the simple model of \citet{appenzeller:2005}, the following
parametrisation of colliminated wind/jet (region 3 in Fig.~\ref{fig:overview})
is adopeted. The wind velocity field, $\mathbf{v}_{\mathrm{wind}}\left(r,\theta\right)$,
is consist of radial and polar components which depend only on $r$
and $\theta$ respectively. The radial componet $v_{r}\left(r\right)$
is assumed to be in the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is asuumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:

>>>>

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,useAMS, usenatbib]{mn2e}
%\usepackage[T1]{fontenc}
%\usepackage[latin1]{inputenc}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amssymb}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{times}

@@

We present  radiative transfer models of classical T~Tauri stars, and investigate the formation of H$\alpha$
in order to understand their complex circumstellar environment, and to explain
the wide variety of emission line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with only magnetospherical accretion to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind) presented,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations, 
although the inclination dependence of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al , and discuss the basic physical conditions required to reproduce
the profiles in each classified type. Using the hybrid model, we fit
the observed H$\alpha$ from the optical component of T~Tau triple
system (T~Tau~N), and find its ratio of the mass-loss to mass-accretion
rate to be $\sim0.18$ with the cold ($T_{\mathrm{wind}}\approx8000\,\mathrm{K}$)
bipolar wind model .

@@

\begin{keywords}
stars: formation -- circumstellar matter -- radiative transfer -- stars: pre-main-sequence
\end{keywords}

@@

T~Tauri stars (TTS) are young ($<\sim3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass stars, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some
of the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit an excess
amount of continuum flux in the ultraviolet (UV) and infrared (IR).
Their spectral energy distribution and polarisation data suggest the
presence of a circumstellar disc, which plays an important role in
regulating dynamics of gas flows around CTTS.

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidence for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet{takami:2003}
show the direct evidence for the presence of bipolar and monopolar
outflows down to $\sim1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observation of high-resolution ($R=50\,000$)
two-dimensional spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc, and in both above and below the disc --- suggesting the presence
of the bipolar outflows. On a larger scale, \emph{HST} observation
of HH30 \citet{burrows:1996} shows the jet traced to within $\lesssim30$~AU
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 AU \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their samples (30 CTTS) show blue-shifted
absorption components in at least one of Balmer lines and Ca~K (most
commonly in H$\alpha$).

In the currently favoured model of accretion flows around CTTS, the
accretion discs are disrupted by the magnetosphere of stars which
channels the gas from the disc onto the surface of the stars (e.g.,
\citealt{uchida:1985}; \citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture of the accretion flows is supported
by the evidences that CTTS have relatively strong ($\sim10^{3}\mathrm{G}$)
magnetic field (e.g., \citealt*{jonhs-krull:1999}; \citealt{guenther:1996})
and by the radiative transfer models which reproduce the observed
profiles for some TTS (\citealt{muzerolle:2001}). The magnetospherical
accretion model naturally explains the blue-ward asymmetric emission
lines (seen in some of CTTS) caused by the partial occultation of
the flow by the disc, and the redshifted absorption component at the
typical free-fall velocities (a few hundred $\mathrm{km\, s^{-1}}$)
seen in some of CTTS. Despite the success of the magnetospherical
accretion model in explaining the line profiles in some CTTS, the
overwhelming observational evidences for the outflow (mentioned above)
in the CTTS profiles suggests that this model is only a part of a
whole picture. Clearly, the modification to include the out-flowing
wind/jet flow is necessary if one requires to predict the mass-accretion
rate and the mass-loss rate of CTTS by modelling their emission profiles
(e.g. H$\alpha$).

Priory to the magnetospherical models, many alternative models had
been considered to explain the observed spectroscopic features mentioned
earlier. For example, (1)~the Alfv\'{e}n wave-driven wind model
(e.g.~\citealt{decampli:1981}; \citealt{hartmann:1982}), (2)~turbulent
boundary layer (between the accretion disc and stellar surface) model
(e.g.~\citealt{bertout:1988}; \citealt{basri:1989}), (3)~chromospheric
model (e.g.~\citealt{calvet:1984}), and (4)~disc wind model (e.g.
\citealt{calvet:1992b}; \citealt{kwan:1995}). Considering the success
of the magnetospherical accretion flow model in some cases and the
observational evidences for the outflows, it is likely that an improved
spectroscopic model requires both inflow and outflow components. The
combination of the magnetospheric accretion flow model with model
1 or 4 would be a reasonable starting point. Although time consuming
to explore a larger parameter space, more realistic density, velocity
and temperature structure from magnetohydrodynamic wind+accretion
models should be considered as inputs for an radiative transfer model
in the future. Recent work by \citet{alencar:2005} showed the observed
H$\alpha$, H$\beta$ and Na~D lines of RW Aur are better produced
by the radiative transfer model which include the collimated disc-wind
arising from near the inner edge of the accretion disc. \citet{malbet:2005}
considered the combination of a spherical wind and the accretion disc
in their radiative transfer model to reproduce Br$\gamma$ from the
early-type Herbig Be star MWC~297.

The magneto-centrifugal wind model, first proposed by \citet{blandford:1982},
has been often used to reproduce the large-scale wind structure of
T Tauri stars, or to model observed optical jets (e.g.~HH~30 jet
by \citealt{burrows:1996}; \citealt{ray:1996}). The launching of
the wind from a Keplerian disc is typically done by treating the equatorial
plane of the disc as a mass-injecting boundary condition (e.g., \citealt{shu:1994};
\citealt{ustyugova:1995}; \citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).
Depending on the location of the open magnetic fields anchored to
the disc, two different types of winds are produced. If the field
is constrained to be near the co-rotation radius of stellar magnetosphere,
an {}``X-wind'' \citep{shu:1994} is produces. If the open field
lines are located in a wider area of the disc, a {}``disc-wind''
similar to \citet{koenigl:2000} is produced \citet{krasnopolsky:2003}.
Recent reviews on the jet/wind-disc connection can be found in \citet{koenigl:2000}
and \citet{pudritz:2005}. Interestingly, \citet{matt:2005} showed
the possibility that the stellar wind along the open magnetic field
originating from the star can cause significant spin-down torque on
the star, provided that mass-loss rate is high enough.

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to perform
the empirical studies of the line formation. More specifically, we
present disc--wind--magnetosphere hybrid models for CTTS, and perform
parametric studies of the H$\alpha$ formation. This study should
provide preliminary physical conditions which lead to the wide variety
of emission line profiles seen in the observation, and would help
to construct more comprehensive circumstellar models (e.g. via MHD
simulations) of T~Tauri stars. the Using the model results, we examine
the H$\alpha$ spectroscopic classification used by \citet{reipurth:1996},
and compare models with observations. We will also discuss whether
our model is consistent with some predictions made by the recent MHD
studies i.e.~ $\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

In section~\ref{sec:model-configuration}, the model assumptions,
and the basic model configurations are presented. We discuss the radiative
transfer model used to compute the profiles in section~\ref{sec:radiative-transfer-model},
and the results of model calculations are given in section~\ref{sec:Results}.
The closer examination and discussion of the results are presented
in section~\ref{sec:Discussion}. Finally, the summary of this work
and the conclusion are in section~\ref{sec:Conclusions}.

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model spaced is divided
into four different regions: 1.~Central continuum source, 2.~magnetosphere
accretion flow, 3.~wind outflow, and 4.~accretion disc. Fig.~\ref{fig:config_dipolar}
depicts the relative location in the model space. In all regions,
the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas freely fall, moving along the magnetic filed onto the
surface of the star. For the purpose of computational simplicity,
the collimated bipolar wind are defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described. In our models, two different
types of the wind configurations are considered: (1)~the bipolar
wind resembling the density structure of the MHD simulation by \citealt{krasnopolsky:2003}
in large scale ($\sim100$~AU) in section~\ref{sub:wind_jet}, and
(2) that in small scale ($\sim10$~AU) in section~\ref{sub:Alternative-model-disc-wind}.

@@

\includegraphics[%
  scale=0.5]{figures/overview3.eps}

@@

\caption{Basic Model Configuration. The system consist of four components: (1)~the continuum source located at the origin $\left( o \right)$ of the cartisian coordinates $\left(x,y,z\right)$ -- the y-axis are into the paper, (2)~magnetospherical accretion flow, (3)~bipolar wind outflow, and (4)~accretion disc. The density distribution is symmetric around the z-axis. The innermost raius of the magnetsphere (at the equotrial plane) coincides with the the inner radius of the accretion disc. From the inner most part of the accretion disc, the gas freely fall, moving along the magneic filed onto the surface of the star. For a computational simplicity, the bipolar wind starts from just outside of the largest radius ($R_{\mathrm{mo}}$) of the magnesophere (dotted line). }

\label{fig:config_dipolar}

@@

\subsection{Continuum Source}

\label{sub:Continuum-Source}

Unless specified otherwise, we adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, R_{\sun}$, $0.5\, M_{\sun}$, and $4000\,\mathrm{K}$
respectively. The model atmosphere of \citet{kurucz:1979} equivalently
\citet{kurucz:1993}  with $T_{\mathrm{ph}}=4000\,\mathrm{K}$ and
$\log g_{*}=3.5$ (cgs) is used to find the photospheric continuum
at H$\alpha$ wavelength. The parameters are summarised in Table~\ref{tab:std_parameters}.

For a model which includes magnetospherical accretion, an additional
source of the continuum source is considered. As the in-falling gas
becomes closer to the stellar surface, it is decelerates in a strong
shock, and heated to $\sim10^{6}\,\mathrm{K}$ (with typical parameters).
The X-ray radiation produced in the shock will be absorbed by the
gas located near, and re-emitted in optical and UV radiation (\citealt{koenigl:1991};
\citealt{hartmann:1994}). This will create hot rings near the stellar
surface where the magnetic field intersects with the surface. For
simplicity, the free-falling kinetic energy is assumed to be thermalized
in the radiating layer, and re-emits as blackbody radiation (with
a single temperature). With the parameters of the magnetosphere and
the star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings, and its temperature is
about $6400\,\mathrm{K}$. If the mass-accretion rate is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$,
the ratio of this accretion luminosity to the photospheric luminosity
is about $0.5$. The continuum emission from the hot rings is taken
into account when computing the line profiles.

@@

\begin{tabular}{lccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$v_{\infty}$&
$b$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{*}\right]$&
$\left[\mathrm{AU}\right]$\tabularnewline
\hline
Standard&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$210$&
$4.0$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

We adopt the magnetospherical accretion flow model of \citet{hartmann:1994},
as done so by \citet{muzerolle:2001} and by \citet{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_dipolar}, the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The former value corresponds to the co-rotation radius
of the accretion disc, and the geometry of the magnetic field/stream
lines is kept constant throughout this paper. The geometry of the
magnetosphere is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form: \begin{equation}
r=R_{\mathrm{m}}\sin^{2}\theta\label{eq:dipole}\end{equation}
 \citep[see][]{ghosh:1977} where $r$, and $\theta$ are coordinates
of the field point ($p$) in Fig.~\ref{fig:config_dipolar} in spherical
coordinates, and $R_{\mathrm{m}}$is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$). The range of $R_{\mathrm{m}}$is
restricted between $R_{\mathrm{mi}}$ and $R_{\mathrm{mo}}$. Using
the field geometry above and the conservation of energy, the velocity
and the density of the accreting gas along the steam line are found
as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by \citet{hartmann:1994}
is simply adopted here. They computed the temperature assuming a volumetric
heating rate which is proportional to $r^{-3}$, and using the energy
balance of the radiative cooling rate of \citet{hartmann:1982} and
the heating rate \citep{hartmann:1994}. Although the temperature
structure of the accretion steam could significantly affect the line
source function, for the purpose of the exploring the general characteristics
of the H$\alpha$ formation, this simple form is a reasonable assumption.
\citet{martin:1996} presented the self-consistent determination of
the thermal structure of the in-flowing gas along the dipole magnetic
field (equation~\ref{eq:dipole}) by solving the heat equation couple
to rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors of the cooling process are bremsstrahlung radiation
and line emission from Ca~II and Mg~II ions. The results of \citet{martin:1996}
qualitatively agrees with that of \citet{hartmann:1994}.

\subsection{Bipolar Wind}

@@

As mentioned earlier, the wind model presented here is to simulate
the collimated MHD disc-wind model of \citet{krasnopolsky:2003} in
large scale ($\sim100$~AU) , and the alternative wind model which
resembles the model of \citet{krasnopolsky:2003} in small scale ($\sim10$~AU)
will be presented later (section~\ref{sub:Alternative-model-disc-wind}).
